---
title: "Age-structure"
author: "Vicente J. Ontiveros"
date: "11/20/2022"
output: ioslides_presentation
---


## Aim of these experiments

- Recover a time-varying infectivity parameter in an age-structured SIR with a simple SIR.
- Develop an expert system that detects structural changes.
  
 <!-- - Random patterns in rates. -->
 <!-- - Recover a SIRD SSA with varying $\iota$. -->

----
**Algorithm**: Moving windows.
----

1. Set up initial values for the parameter.
2. For a given window length, estimate the parameters of the model.
3. Set up the parameters as the new initial values.
4. Move the window a given step length. 
5. Repeat 2-4 until a desired condition. 


## Let's have a look at the data 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
```

```{r Load packages and source files}
library(tidyverse)
library(deSolve)
# library(dplyr)
library(lhs)
# library(ggplot2)
# library(tibble)
# library(utils)
library(lubridate)
# library(tidyr)
library(strucchange)

source("ll_function_v6.R")
source("R/fitting_procedure_v5.R")
source('compilation.R')

dyn.load("src/SIR.so")
```



```{r Plot data}
data <- read.csv("data/artificial221115_Age.csv", na.strings = "", fileEncoding = "UTF-8-BOM")
# data <- read.csv("artificial220427.csv", na.strings = "", fileEncoding = "UTF-8-BOM")

# unique(data$countriesAndTerritories)

selected_country <- c("SIRDageComplexS0_SSA-s0_N10000", 
                      "SIRDageComplexS1_SSA-s0_N10000",
                      "SIRDageComplexS2_SSA-s0_N10000",
                      "SIRDageComplexS3_SSA-s0_N10000")

 country <-
  data %>% filter(countriesAndTerritories %in% selected_country) %>%
  select(dateRep, cases, deaths, countriesAndTerritories, popData2020) %>%
  mutate(date_new = dmy(dateRep)) %>% arrange(date_new)

ggplot(country %>% filter(date_new < dmy("23022020") + 120), aes(x = date_new)) +
  geom_ribbon(aes(ymax = cases, ymin = 0), fill = "darkgreen", alpha = .8) +
  geom_ribbon(aes(ymax = deaths, ymin = 0), fill = "purple", alpha = .8) +
  facet_wrap(~ countriesAndTerritories) +
  theme_bw() + theme(panel.grid.minor = element_blank(), aspect.ratio = .618) +
  ylab("Daily cases & deaths") + xlab("Date")
```

S{0, 1, 2, 3} differ on the length of the reduction in infectivity.

## First try -- S0_SSA-s0_N10000

```{r SIRDageComplexS0_SSA-s0_N10000}
selected_country <- "SIRDageComplexS0_SSA-s0_N10000"

sims <- 3 #number of "threads" to be optimized
days <- 70 #number of jumps of the timestep

set.seed(476468713)


#### Trying to estimate lockdown with 7d data

seq_pars <- data.frame(iota = rep(NA, sims*days), day = rep(NA, sims*days))

# country
  
  
country <-
  data %>% filter(countriesAndTerritories == selected_country) %>% 
  select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
  separate(infected, sep = ",", letters[1:3]) %>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         infected = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
  separate(recovered, sep = ",", letters[1:3]) %>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         recovered = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
  separate(dead, sep = ",", letters[1:3])%>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         dead = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% select(-letters[1:3]) %>% 
  select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
  mutate(date_new = dmy(dateRep)) %>% arrange(date_new)

seeds <- randomLHS(sims, 1)

seeds <- t(seeds) * c(.5)

#### Fitting

pop <- country$popData2020[1]

start_0 <- as.Date("2020-02-24")

for(i in 0:(days - 1)){
  start <- start_0 + i
end <- start + 6


  n_start <- which(country$date_new == start)
  n_end <- which(country$date_new == end)
  input <- country[n_start:n_end, ]
 
  
  devs <- c()
spl <- (input$cases)
fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 4)
devs <- sd(spl - predict(fit)$y)

if(devs < .5) devs <- 1

temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, rho = .03, i_0 = country$infected[n_start - 1], r_0 = country$recovered[n_start - 1] + country$dead[n_start - 1])

temp <- as.data.frame(temp)
colnames(temp) <- c("iota")

seq_pars[(1 + sims * i):(sims + sims * i), ] <-  
                  temp %>% add_column(day = i)

rm(temp)

}

lockdown7d <- seq_pars
# save(lockdown7d, file = "lockdown7d.RData")
ggplot(lockdown7d %>% unique(), aes(x = day, y = iota)) +
  # geom_boxplot(outlier.shape = NA)
  geom_line() + theme_bw() +
  theme(aspect.ratio = .618, panel.grid.minor = element_blank())
```

Daily re-parametrization. Window length = 7d.


## Four scenarios - seed 0

```{r Four scenarios}

# I keep s0, for the 4 scenarios and 4 population sized
# unique(data$countriesAndTerritories)
# 
# c((c(0:3) * 2 + 1), (c(0:3) * 2 + 33), (c(0:3) * 2 + 65))

locks <- unique(data$countriesAndTerritories)[c((c(0:3) * 2 + 1), (c(0:3) * 2 + 33), (c(0:3) * 2 + 65))]

sims <- 3 #number of "threads" to be optimized
days <- 60

seq_pars <- data.frame(iota = rep(NA, sims*days*length(locks)), 
                       day = rep(NA, sims*days*length(locks)), 
                       model = rep(NA, sims*days*length(locks)))
seeds <- randomLHS(sims, 1)
  
  seeds <- t(seeds) * c(.5)
  n <- 0
  for(x in 1:length(locks)){
    method <- locks[x] #There are many options in here. Keep one
    selected_country <- method
    
    country <-
      data %>% filter(countriesAndTerritories == selected_country) %>% 
      select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
      separate(infected, sep = ",", letters[1:3]) %>% 
      mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
             infected = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
      separate(recovered, sep = ",", letters[1:3]) %>% 
      mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
             recovered = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
      separate(dead, sep = ",", letters[1:3])%>% 
      mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
             dead = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% select(-letters[1:3]) %>% 
      select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
      mutate(date_new = dmy(dateRep)) %>% arrange(date_new)
    
    #### Fitting
    
    pop <- country$popData2020[1]
    
    start_0 <- as.Date("2020-02-24")
    
    for(i in 0:(days - 1)){
      start <- start_0 + i
      end <- start + 6
      
      
      n_start <- which(country$date_new == start)
      n_end <- which(country$date_new == end)
      input <- country[n_start:n_end, ]
      
      
      devs <- c()
      spl <- (input$cases)
      fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 4)
      devs <- sd(spl - predict(fit)$y)
      
      if(devs < .5) devs <- 1
      
      temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, rho = .03, i_0 = country$infected[n_start - 1], r_0 = country$recovered[n_start - 1] + country$dead[n_start - 1])
      
      temp <- as.data.frame(temp)
      colnames(temp) <- c("iota")
      
      seq_pars[(1 + sims * i + n * sims * days):(sims + sims * i + n * sims * days), ] <-  
        temp %>% add_column(day = i) %>% add_column(model = method)
      
      rm(temp)
      
    }
    n <- n + 1
  }


scenarios <- inner_join(seq_pars, 
seq_pars %>% select(model) %>% unique() %>% 
  add_column(S = paste0("S", rep(0:3, 3))) %>% 
  add_column(N = paste0("N", rep(c(1000, 10000, 100000), each = 4))))

ggplot(scenarios, aes(day, iota, color = N)) + geom_line() +
  facet_wrap(~ S) + theme_bw() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

Variability decreases with N. Remember $\iota = .2$

## Structural breaks in infectivity. 

_Chow test_. 

We split the data in two groups:

$$
y_t=a_1+b_1x_{1t} + c_1x_{2t} + \varepsilon \, 
$$

$$
y_t=a_2+b_2x_{1t} + c_2x_{2t} + \varepsilon
$$

So, our null hypothesis is $a_1=a_2$, $b_1=b_2$, and $c_1=c_2$.

The test statistic follows an F-distribution with $N - 2k$ d.o.f.     

## Example -- S0_SSA-s0_N10000

```{r}

statistic <- rep(NA, 70)
data <- lockdown7d %>% group_by(day) %>% summarise(iota = mean(iota)) %>% as.data.frame()

for(i in 7:64){
  statistic[i] <- (sctest(data$iota ~ data$day,
               type = 'Chow', point = i))$statistic
}


ggplot(data %>% add_column(statistic), aes(x = day, y = iota)) +
  geom_point() +
  geom_line(aes(y = statistic/1700), color = 'red') + theme_bw()

```

Red line -- statistic.

## Parametrizing the break

```{r break included}

# source("R/fitting_procedure_break_v1.R")
# source("ll_function_break_v1.R")
# 
data <- read.csv("data/artificial221115_Age.csv", na.strings = "", fileEncoding = "UTF-8-BOM")
# 
# selected_country <- "SIRDageComplexS0_SSA-s0_N10000"
# 
# country <-
#   data %>% filter(countriesAndTerritories == selected_country) %>% 
#   select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
#   separate(infected, sep = ",", letters[1:3]) %>% 
#   mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
#          infected = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
#   separate(recovered, sep = ",", letters[1:3]) %>% 
#   mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
#          recovered = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
#   separate(dead, sep = ",", letters[1:3])%>% 
#   mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
#          dead = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% select(-letters[1:3]) %>% 
#   select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
#   mutate(date_new = dmy(dateRep)) %>% arrange(date_new)
# 
# input <- country[2:71, ]
# input <- input %>% select(cases)
#   
#   devs <- c()
# spl <- (input$cases)
# fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 4)
# devs <- sd(spl - predict(fit)$y)
# 
# if(devs < .5) devs <- 1
# 
# seeds <- randomLHS(100, 3)
#   
#   seeds <- t(seeds) * c(.5, .5, 50)
# 
# temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, rho = .03, i_0 = country$infected[1], r_0 = 0)
# 
# temp <- as.data.frame(temp)
# colnames(temp) <- c("iota_0", 'iota_1', 'break')
# save(temp, file = "data/temp_break.RData")

# For some reason, this chunk works manually but not when knitting. I save and
# load the results.

load("data/temp_break.RData")

ggplot(temp %>% pivot_longer(cols = 1:3), aes(y = value)) +
  geom_boxplot() +
  facet_wrap(~ name, scales = 'free') + 
  theme_bw() +
  theme(panel.grid.minor = element_blank(), axis.ticks.x = element_blank(),
        axis.text.x = element_blank(), panel.grid.major.x = element_blank())



```

The break is recovered nicely.

## Effect of window length

```{r}
selected_country <- "SIRDageComplexS0_SSA-s0_N10000"
country <-
  data %>% filter(countriesAndTerritories == selected_country) %>% 
  select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
  separate(infected, sep = ",", letters[1:3]) %>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         infected = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
  separate(recovered, sep = ",", letters[1:3]) %>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         recovered = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
  separate(dead, sep = ",", letters[1:3])%>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         dead = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% select(-letters[1:3]) %>% 
  select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
  mutate(date_new = dmy(dateRep)) %>% arrange(date_new)

sims <- 3 #number of "threads" to be optimized
days <- 70 #number of jumps of the timestep

seeds <- randomLHS(sims, 1)
seeds <- t(seeds) * c(.5)
set.seed(476468713)
pop <- country$popData2020[1]

wl.df <- data.frame()

for(wl in 5:21){
  seq_pars <- data.frame(iota = rep(NA, sims*days), day = rep(NA, sims*days))
  
  #### Fitting
  
  start_0 <- as.Date("2020-02-24")
  
  for(i in 0:(days - 1)){
    start <- start_0 + i
    end <- start + wl - 1
    
    n_start <- which(country$date_new == start)
    n_end <- which(country$date_new == end)
    input <- country[n_start:n_end, ]
    
    devs <- c()
    spl <- (input$cases)
    fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 4)
    devs <- sd(spl - predict(fit)$y)
    
    if(devs < .5) devs <- 1
    
    temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, rho = .03, i_0 = country$infected[n_start - 1], r_0 = country$recovered[n_start - 1] + country$dead[n_start - 1])
    
    temp <- as.data.frame(temp)
    colnames(temp) <- c("iota")
    
    seq_pars[(1 + sims * i):(sims + sims * i), ] <-  
      temp %>% add_column(day = i)
    
    
    rm(temp)
    
  }
seq_pars <- seq_pars %>% add_column(Window.length = wl)
  wl.df <- rbind(wl.df, seq_pars)
}

wl.df <- wl.df %>% unique()

ggplot(wl.df, aes(x = day, y = iota, group = Window.length, color = Window.length)) + 
  geom_line() + theme_bw() + theme(panel.grid.minor = element_blank())

```

## Chow-test

```{r}

chow.df <- data.frame()

for(wl in 5:21){
  statistic <- rep(NA, 70)
data <- wl.df %>% filter(Window.length == wl) %>% group_by(day) %>% summarise(iota = mean(iota)) %>% as.data.frame()

for(i in 7:64){
  statistic[i] <- (sctest(data$iota ~ data$day,
               type = 'Chow', point = i))$statistic
}

temp <- data %>% add_column(statistic) %>% add_column(wl = wl)
chow.df <- rbind(chow.df, temp)
}



ggplot(chow.df, aes(x = day, y = statistic, group = wl, color = wl)) +
  geom_line() + theme_bw()
```

On the infectivities.

## Findings so far

- For this simple setting, the moving window suffices to find the structural change. 
- Incorporating the breakpoint in parameter estimation also helps us find the structural change.
- Chow tests in the infectivity series do not give exact information on when the change happened. 


## What about smaller drops?

```{r}
data <- read.csv("data/artificial230207_Age2.csv", na.strings = "", fileEncoding = "UTF-8-BOM")

# unique(data$countriesAndTerritories)
selected_country <- "SIRDageComplex-IotaScale0.5-S0_SSA-s0_N10000"

sims <- 3 #number of "threads" to be optimized
days <- 70 #number of jumps of the timestep

set.seed(476468713)


#### Trying to estimate lockdown with 7d data

seq_pars <- data.frame(iota = rep(NA, sims*days), day = rep(NA, sims*days))

# country
  
  
country <-
  data %>% filter(countriesAndTerritories == selected_country) %>% 
  select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
  separate(infected, sep = ",", letters[1:3]) %>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         infected = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
  separate(recovered, sep = ",", letters[1:3]) %>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         recovered = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
  separate(dead, sep = ",", letters[1:3])%>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         dead = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% select(-letters[1:3]) %>% 
  select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
  mutate(date_new = dmy(dateRep)) %>% arrange(date_new)

seeds <- randomLHS(sims, 1)

seeds <- t(seeds) * c(.5)

#### Fitting

pop <- country$popData2020[1]

start_0 <- as.Date("2020-02-24")

for(i in 0:(days - 1)){
  start <- start_0 + i
end <- start + 6


  n_start <- which(country$date_new == start)
  n_end <- which(country$date_new == end)
  input <- country[n_start:n_end, ]
 
  
  devs <- c()
spl <- (input$cases)
fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 4)
devs <- sd(spl - predict(fit)$y)

if(devs < .5) devs <- 1

temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, rho = .03, i_0 = country$infected[n_start - 1], r_0 = country$recovered[n_start - 1] + country$dead[n_start - 1])

temp <- as.data.frame(temp)
colnames(temp) <- c("iota")

seq_pars[(1 + sims * i):(sims + sims * i), ] <-  
                  temp %>% add_column(day = i)

rm(temp)

}

lockdown7d <- seq_pars
# save(lockdown7d, file = "lockdown7d.RData")
ggplot(lockdown7d %>% unique(), aes(x = day, y = iota)) +
  # geom_boxplot(outlier.shape = NA)
  geom_line() + theme_bw() +
  theme(aspect.ratio = .618, panel.grid.minor = element_blank()) +
  scale_y_continuous(limits = c(0, NA))
```

Let's try with a iota reduction of .5, 7d window.

## Window length reveals lockdown
```{r}
wl.df <- data.frame()

for(wl in 5:21){
  seq_pars <- data.frame(iota = rep(NA, sims*days), day = rep(NA, sims*days))
  
  #### Fitting
  
  start_0 <- as.Date("2020-02-24")
  
  for(i in 0:(days - 1)){
    start <- start_0 + i
    end <- start + wl - 1
    
    n_start <- which(country$date_new == start)
    n_end <- which(country$date_new == end)
    input <- country[n_start:n_end, ]
    
    devs <- c()
    spl <- (input$cases)
    fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 4)
    devs <- sd(spl - predict(fit)$y)
    
    if(devs < .5) devs <- 1
    
    temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, rho = .03, i_0 = country$infected[n_start - 1], r_0 = country$recovered[n_start - 1] + country$dead[n_start - 1])
    
    temp <- as.data.frame(temp)
    colnames(temp) <- c("iota")
    
    seq_pars[(1 + sims * i):(sims + sims * i), ] <-  
      temp %>% add_column(day = i)
    
    
    rm(temp)
    
  }
seq_pars <- seq_pars %>% add_column(Window.length = wl)
  wl.df <- rbind(wl.df, seq_pars)
}

wl.df <- wl.df %>% unique()

ggplot(wl.df, aes(x = day, y = iota, group = Window.length, color = Window.length)) + 
  geom_line() + theme_bw() + theme(panel.grid.minor = element_blank()) +
  geom_vline(xintercept = 30)
```

The lockdown started on day 30. Iota factor .5

## Let's see with lesser drops (I)

```{r}
data <- read.csv("data/artificial230207_Age2.csv", na.strings = "", fileEncoding = "UTF-8-BOM")

selected_country <- "SIRDageComplex-IotaScale0.75-S0_SSA-s0_N10000"

sims <- 3 #number of "threads" to be optimized
days <- 70 #number of jumps of the timestep

set.seed(476468713)


#### Trying to estimate lockdown with 7d data

seq_pars <- data.frame(iota = rep(NA, sims*days), day = rep(NA, sims*days))

# country
  
  
country <-
  data %>% filter(countriesAndTerritories == selected_country) %>% 
  select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
  separate(infected, sep = ",", letters[1:3]) %>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         infected = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
  separate(recovered, sep = ",", letters[1:3]) %>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         recovered = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
  separate(dead, sep = ",", letters[1:3])%>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         dead = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% select(-letters[1:3]) %>% 
  select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
  mutate(date_new = dmy(dateRep)) %>% arrange(date_new)

seeds <- randomLHS(sims, 1)

seeds <- t(seeds) * c(.5)

wl.df <- data.frame()

for(wl in 5:21){
  seq_pars <- data.frame(iota = rep(NA, sims*days), day = rep(NA, sims*days))
  
  #### Fitting
  
  start_0 <- as.Date("2020-02-24")
  
  for(i in 0:(days - 1)){
    start <- start_0 + i
    end <- start + wl - 1
    
    n_start <- which(country$date_new == start)
    n_end <- which(country$date_new == end)
    input <- country[n_start:n_end, ]
    
    devs <- c()
    spl <- (input$cases)
    fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 4)
    devs <- sd(spl - predict(fit)$y)
    
    if(devs < .5) devs <- 1
    
    temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, rho = .03, i_0 = country$infected[n_start - 1], r_0 = country$recovered[n_start - 1] + country$dead[n_start - 1])
    
    temp <- as.data.frame(temp)
    colnames(temp) <- c("iota")
    
    seq_pars[(1 + sims * i):(sims + sims * i), ] <-  
      temp %>% add_column(day = i)
    
    
    rm(temp)
    
  }
seq_pars <- seq_pars %>% add_column(Window.length = wl)
  wl.df <- rbind(wl.df, seq_pars)
}

wl.df <- wl.df %>% unique()

ggplot(wl.df, aes(x = day, y = iota, group = Window.length, color = Window.length)) + 
  geom_line() + theme_bw() + theme(panel.grid.minor = element_blank()) +
  geom_vline(xintercept = 30)

```

There's a bit more variance now. Factor .75


## Let's see with lesser drops (II)

```{r}
data <- read.csv("data/artificial230207_Age2.csv", na.strings = "", fileEncoding = "UTF-8-BOM")

selected_country <- "SIRDageComplex-IotaScale0.875-S0_SSA-s0_N10000"

sims <- 3 #number of "threads" to be optimized
days <- 100 #number of jumps of the timestep

set.seed(476468713)


#### Trying to estimate lockdown with 7d data

seq_pars <- data.frame(iota = rep(NA, sims*days), day = rep(NA, sims*days))

# country
  
  
country <-
  data %>% filter(countriesAndTerritories == selected_country) %>% 
  select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
  separate(infected, sep = ",", letters[1:3]) %>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         infected = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
  separate(recovered, sep = ",", letters[1:3]) %>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         recovered = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% 
  separate(dead, sep = ",", letters[1:3])%>% 
  mutate(a = str_sub(a, start = 2), c = str_sub(c, end = -2), 
         dead = as.numeric(a) + as.numeric(b) + as.numeric(c)) %>% select(-letters[1:3]) %>% 
  select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
  mutate(date_new = dmy(dateRep)) %>% arrange(date_new)

seeds <- randomLHS(sims, 1)

seeds <- t(seeds) * c(.5)

wl.df <- data.frame()

for(wl in 5:21){
  seq_pars <- data.frame(iota = rep(NA, sims*days), day = rep(NA, sims*days))
  
  #### Fitting
  
  start_0 <- as.Date("2020-02-24")
  
  for(i in 0:(days - 1)){
    start <- start_0 + i
    end <- start + wl - 1
    
    n_start <- which(country$date_new == start)
    n_end <- which(country$date_new == end)
    input <- country[n_start:n_end, ]
    
    devs <- c()
    spl <- (input$cases)
    fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 4)
    devs <- sd(spl - predict(fit)$y)
    
    if(devs < .5) devs <- 1
    
    temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, rho = .03, i_0 = country$infected[n_start - 1], r_0 = country$recovered[n_start - 1] + country$dead[n_start - 1])
    
    temp <- as.data.frame(temp)
    colnames(temp) <- c("iota")
    
    seq_pars[(1 + sims * i):(sims + sims * i), ] <-  
      temp %>% add_column(day = i)
    
    
    rm(temp)
    
  }
seq_pars <- seq_pars %>% add_column(Window.length = wl)
  wl.df <- rbind(wl.df, seq_pars)
}

wl.df <- wl.df %>% unique()

ggplot(wl.df, aes(x = day, y = iota, group = Window.length, color = Window.length)) + 
  geom_line() + theme_bw() + theme(panel.grid.minor = element_blank()) +
  geom_vline(xintercept = 30)

```

It's more difficult to see the exact day of the lockdown. Factor .875

## Chow once more

```{r}

days <- 70

statistic <- rep(NA, days)
pval <- rep(NA, days)
data <- wl.df %>% group_by(day) %>% summarise(sd = sd(iota)) %>% slice(1:days) %>%  as.data.frame()

for(i in 7:(days-6)){
  statistic[i + 1] <- (sctest(data$sd ~ data$day,
               type = 'Chow', point = i))$statistic
  pval[i + 1] <- (sctest(data$sd ~ data$day,
               type = 'Chow', point = i))$p.value
}


ggplot(data %>% add_column(statistic), aes(x = day, y = sd)) +
  geom_point() +
  geom_line(aes(y = statistic/3256), color = 'red') + theme_bw()

# ggplot(wl.df %>% group_by(day) %>% summarise(sd = sd(iota)), aes(x = day, y = sd)) +
#   geom_line() + theme_bw() + theme(panel.grid.minor = element_blank()) +
#   geom_vline(xintercept = 30)
```

Red line - Chow test statistic. Points, sd of the window estimates. 

## Chow test again

```{r}
days <- 100

statistic <- rep(NA, days)
pval <- rep(NA, days)
data <- wl.df %>% group_by(day) %>% summarise(sd = sd(iota)) %>% slice(1:days) %>% as.data.frame()

for(i in 7:(days-6)){
  statistic[i + 1] <- (sctest(data$sd ~ data$day,
               type = 'Chow', point = i))$statistic
  pval[i + 1] <- (sctest(data$sd ~ data$day,
               type = 'Chow', point = i))$p.value
}


ggplot(data %>% add_column(statistic), aes(x = day, y = sd)) +
  geom_point() +
  geom_line(aes(y = statistic/3256), color = 'red') + theme_bw()

# ggplot(wl.df %>% group_by(day) %>% summarise(sd = sd(iota)), aes(x = day, y = sd)) +
#   geom_line() + theme_bw() + theme(panel.grid.minor = element_blank()) +
#   geom_vline(xintercept = 30)
```

Red line - Chow test statistic. Points, sd of the window estimates. 


## More open questions
- What about goodness of fit/prediction?
- More unknown parameters?

## Next step

- Evaluate goodness of fit/prediction. 
- Now, parameters are equal for all ages. What if we have parameters that are not equal and changing?

