---
title: "Fitting artificial data - Lockdown"
output: html_document
---

#### Loading packages and source files.

```{r}
library(deSolve)
library(dplyr)
library(lhs)
library(ggplot2)
library(tibble)
library(utils)
library(lubridate)
library(tidyr)


source("ll_function_v6.R")
source("fitting_procedure_v4.R")

dyn.load("SIR.so")
```

## Artificial data - Lockdown

## Fitting.

Here, I try to fit a SIR model to a SIRD stochastic with a lockdown. I start easy.

#### Number of threads

```{r}

sims <- 100 #number of "threads" to be optimized

```

#### Setting seed



```{r}

set.seed(476468713)

```





#### Trying to estimate lockdown with 7d data

```{r}
data <- read.csv("artificial220427.csv", na.strings = "", fileEncoding = "UTF-8-BOM")


seq_pars <- data.frame(iota = rep(NA, 150 * 100), day = rep(NA, 150 * 100))


method <- "SIRDlockdown12a_SSA-s0_N100000" #There are many options in here. Keep one
  selected_country <- method

country <-
  data %>% filter(countriesAndTerritories == selected_country) %>% 
  select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
  mutate(infected = as.numeric(infected), recovered = as.numeric(recovered), 
         dead = as.numeric(dead), date_new = dmy(dateRep)) %>% arrange(date_new)





seeds <- randomLHS(sims, 1)

seeds <- t(seeds) * c(.5)

#### Fitting

pop <- country$popData2020[1]

start_0 <- as.Date("2020-02-24")

for(i in 0:149){
  start <- start_0 + i
end <- start + 6


  n_start <- which(country$date_new == start)
  n_end <- which(country$date_new == end)
  input <- country[n_start:n_end, ]
 
  
  devs <- c()
spl <- (input$cases)
fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 4)
devs <- sd(spl - predict(fit)$y)


temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, rho = .03, i_0 = country$infected[n_start - 1], r_0 = country$recovered[n_start - 1] + country$dead[n_start - 1])

temp <- as.data.frame(temp)
colnames(temp) <- c("iota")

seq_pars[(1 + 100 * i):(100 + 100 * i), ] <-  
                  temp %>% add_column(day = i)

rm(temp)

}
```


## Plotting parameters

Remember that the true values for $\iota$ was $.2$.


```{r}
lockdown7d <- seq_pars
save(lockdown7d, file = "lockdown7d.RData")
ggplot(lockdown7d, aes(x = day, y = iota, group = day)) + geom_boxplot(outlier.shape = NA) 
```


#### Trying to estimate lockdown with 14d data

```{r}
data <- read.csv("artificial220427.csv", na.strings = "", fileEncoding = "UTF-8-BOM")

sims <- 30 #Optimization for a single variable produce a global optimum, it seems.
interval <- 14 #Days used for optimization

seq_pars <- data.frame(iota = rep(NA, 150 * sims), day = rep(NA, 150 * sims))


method <- "SIRDlockdown12a_SSA-s0_N100000" #There are many options in here. Keep one
  selected_country <- method

country <-
  data %>% filter(countriesAndTerritories == selected_country) %>% 
  select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
  mutate(infected = as.numeric(infected), recovered = as.numeric(recovered), 
         dead = as.numeric(dead), date_new = dmy(dateRep)) %>% arrange(date_new)





seeds <- randomLHS(sims, 1)

seeds <- t(seeds) * c(.5)

#### Fitting

pop <- country$popData2020[1]

start_0 <- as.Date("2020-02-24")

for(i in 0:149){
  start <- start_0 + i
end <- start + interval - 1


  n_start <- which(country$date_new == start)
  n_end <- which(country$date_new == end)
  input <- country[n_start:n_end, ]
 
  
  devs <- c()
spl <- (input$cases)
fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 4)
devs <- sd(spl - predict(fit)$y)


temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, rho = .03, i_0 = country$infected[n_start - 1], r_0 = country$recovered[n_start - 1] + country$dead[n_start - 1])

temp <- as.data.frame(temp)
colnames(temp) <- c("iota")

seq_pars[(1 + sims * i):(sims + sims * i), ] <-  
                  temp %>% add_column(day = i)

rm(temp)

}
```


## Plotting parameters

Remember that the true values for $\iota$ was $.2$.


```{r}
lockdown14d <- seq_pars
save(lockdown14d, file = "lockdown14d.RData")
ggplot(lockdown14d, aes(x = day, y = iota, group = day)) + geom_boxplot(outlier.shape = NA) 
```

#### Plotting both at the same time

```{r}
lockdown <- rbind(
lockdown7d %>% group_by(day) %>% summarise(iota = mean(iota)) %>% add_column(Interval = "7d"),
lockdown14d %>% group_by(day) %>% summarise(iota = mean(iota)) %>% add_column(Interval = "14d"))

ggplot(lockdown, aes(day, iota, color = Interval)) + geom_line()
```



## Fit now all lockdown models


```{r}
data <- read.csv("artificial220510.csv", na.strings = "", fileEncoding = "UTF-8-BOM")

unique(data$countriesAndTerritories)
locks <- unique(data$countriesAndTerritories)[c(9, 11, 13, 77, 79, 81, 145, 147, 149)]

sims <- 30 #Optimization for a single variable produce a global optimum, it seems.
interval <- 7 #Days used for optimization

seq_pars <- data.frame(iota = rep(NA, 150 * length(locks)), 
                       day = rep(NA, 150 * length(locks)), 
                       model = rep(NA, 150 * length(locks)))
seeds <- randomLHS(sims, 1)
  
  seeds <- t(seeds) * c(.5)

  
  # There's some error here. Had to run it by hand.
  
#   for(x in 1:length(locks)){
#   method <- locks[x] #There are many options in here. Keep one
#   selected_country <- method
#   
#   country <-
#     data %>% filter(countriesAndTerritories == selected_country) %>% 
#     select(dateRep, cases, countriesAndTerritories, popData2020, infected, recovered, dead) %>% 
#     mutate(infected = as.numeric(infected), recovered = as.numeric(recovered), 
#            dead = as.numeric(dead), date_new = dmy(dateRep)) %>% arrange(date_new)
#   
#   #### Fitting
#   
#   pop <- country$popData2020[1]
#   
#   start_0 <- as.Date("2020-02-24")
#   
#   for(i in 0:149){
#     start <- start_0 + i
#     end <- start + interval - 1
#     
#     
#     n_start <- which(country$date_new == start)
#     n_end <- which(country$date_new == end)
#     input <- country[n_start:n_end, ]
#     
#     
#     devs <- c()
#     spl <- (input$cases)
#     fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 4)
#     devs <- sd(spl - predict(fit)$y)
#     
#     
#     temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, rho = .03, i_0 = country$infected[n_start - 1], r_0 = country$recovered[n_start - 1] + country$dead[n_start - 1])
#     
#     temp <- as.data.frame(temp)
#     colnames(temp) <- c("iota")
#   
#     seq_pars[(1 + i + (x - 1) * 150), ] <-  
#       temp %>% slice(1) %>% add_column(day = i) %>% add_column(model = locks[x])
#     
#     rm(temp)
#     
#   }
# }

```


```{r}
load("lockdowns.RData")

ggplot(seq_pars, aes(day, iota, color = model)) + geom_line() + 
  scale_y_continuous(limits = c(0, .5)) + theme(legend.position = "bottom")
```


