---
title: "Fitting artificial data - random patterns in infection rate"
author: Vicente J. Ontiveros
output: ioslides_presentation
---

## Aim of these experiments

 - Examine whether we are able to correctly recover the parameters of different models with a simple SIR. 
 - Random patterns in rates.
 - Recover a SIRD SSA with varying $\iota$.
 
   1. Different seeds.
   2. Population size.
   3. Intensity of fluctuations.
   
## Some data
```{r setup, include=FALSE}
Sys.time()
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
```

```{r Load packages and source files}
library(deSolve)
library(dplyr)
library(lhs)
library(ggplot2)
library(tibble)
library(utils)
library(lubridate)
library(tidyr)


source("ll_function_v5.R")
source("fitting_procedure_v3.R")

dyn.load("SIR.so")
```


```{r Plot data}
data <- read.csv("artificial220517.csv", na.strings = "", fileEncoding = "UTF-8-BOM")
# data <- read.csv("artificial220427.csv", na.strings = "", fileEncoding = "UTF-8-BOM")

selected_country <- "SIRDvaryingIota-NormalSD0.1_SSA-s0_N10000" 
# selected_country <- "SIRD_ODE_N10000" 
# nrow(data %>% unique())   
country <-
  data %>% unique() %>% filter(countriesAndTerritories == selected_country) %>%
  select(dateRep, cases, deaths, countriesAndTerritories, popData2020) %>%
  mutate(date_new = dmy(dateRep)) %>% arrange(date_new)

ggplot(country %>% filter(date_new < dmy("23022020") + 120), aes(x = date_new)) +
  geom_ribbon(aes(ymax = cases, ymin = 0), fill = "darkgreen", alpha = .8) +
  geom_ribbon(aes(ymax = deaths, ymin = 0), fill = "purple", alpha = .8) +
  theme_bw() + theme(panel.grid.minor = element_blank(), aspect.ratio = .618) +
  ylab("Daily cases & deaths") + xlab("Date")
```

```{r}
cat(selected_country)
```

## First try -- SD0.1 SSA-s0 N10000

```{r SD0.1 SSA-s0 N10000}
country <- country %>% slice(-1)

cum_cases <- country %>% mutate(cum_cases = cumsum(cases))

last20 <- country %>% select(date_new, cases) %>% 
  filter(date_new > "2020-02-22") %>%
  mutate(lag1 = lag(cases, 1), lag2 = lag(cases, 2), 
                          lag3 = lag(cases, 3), lag4 = lag(cases, 4), 
                          lag5 = lag(cases, 5), lag6 = lag(cases, 6), 
                          lag7 = lag(cases, 7), lag8 = lag(cases, 8), 
                          lag9 = lag(cases, 9), lag10 = lag(cases, 10), 
                          lag11 = lag(cases, 11), lag12 = lag(cases, 12), 
                          lag13 = lag(cases, 13), lag14 = lag(cases, 14), 
                          lag15 = lag(cases, 15), lag16 = lag(cases, 16), 
                          lag17 = lag(cases, 17), lag18 = lag(cases, 18), 
                          lag19 = lag(cases, 19), lag20 = lag(cases, 20)) %>% 
  mutate_at(3:22, replace_na, 0) %>% 
mutate(last20 = (cases + lag1 + lag2 + lag3 + lag4 + lag5 + lag6 + lag7 + lag8 + lag9 + lag10 + lag11 + lag12 + lag13 + lag14 + lag15 + lag16 + lag17 + lag18 + lag19 + lag20)) %>% select(date_new, last20) #%>% 


sims <- 100

seeds <- randomLHS(sims, 4)

seeds <- t(seeds) * c(.5, .1, 6, 1)

#### Fitting

pop <- country$popData2020[1]

start <- as.Date("2020-02-24")
end <- start + 120


  n_start <- which(country$date_new == start)
  n_end <- which(country$date_new == end)
  input <- country[n_start:n_end, ]
 
  
  devs <- c()
spl <- (input$cases)
fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 15)
devs <- sd(spl - predict(fit)$y)

# plot(fit, ylim = c(0,500))

temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, i_lim = 20, r_lim = 10)

temp <- as.data.frame(temp)
colnames(temp) <- c("iota", "rho", "Initial infected", "Initial recovered")

pars <- temp %>% add_column(Method = selected_country) %>% pivot_longer(!Method, names_to = 'par', values_to = 'value')

ggplot(pars, aes(x = value, color = Method)) + geom_freqpoly(bins = 100) +
  facet_wrap(~ par, scales = "free")  +
  theme_bw() + 
  theme(legend.position = "none", aspect.ratio = .618, panel.grid.minor = element_blank())
# seq_pars <- rbind(seq_pars, 
#                   temp %>% add_column(Method = method))
# 
rm(temp)

# ggplot(pars, aes(x = Method, y = value, groups = Method)) + geom_boxplot(outlier.shape = NA) + 
#   facet_wrap(~ par, scales = "free_y") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

Remember that $\iota = .2$, $\rho = .02$, and $I_0 = 3$.

```{r}
save(pars, file = "SD0.1_SSA-s0_N10000.RData")
```

# Different seeds

## N10000 SD0.1
```{r}
models <- c("SIRDvaryingIota-NormalSD0.1_SSA-s0_N10000",
  "SIRDvaryingIota-NormalSD0.1_SSA-s1_N10000",
            "SIRDvaryingIota-NormalSD0.1_SSA-s2_N10000",
  "SIRDvaryingIota-NormalSD0.1_SSA-s3_N10000",
  "SIRD_ODE_N10000")

legend <- c("s0", "s1", "s2", "s3", "sd = 0")

result <- vector("list", length(models))

for (m in 1:length(models)){
  selected_country <- models[m] 
   
country <-
  data %>% unique() %>% filter(countriesAndTerritories == selected_country) %>%
  select(dateRep, cases, deaths, countriesAndTerritories, popData2020, infected, recovered, dead) %>%
  mutate(infected = as.numeric(infected), recovered = as.numeric(recovered),
         dead = as.numeric(dead), date_new = dmy(dateRep)) %>% arrange(date_new)


country <- country %>% slice(-1)


seeds <- randomLHS(sims, 4)

seeds <- t(seeds) * c(.5, .1, 6, 1)

#### Fitting

end <- start + 120


  n_start <- which(country$date_new == start)
  n_end <- which(country$date_new == end)
  input <- country[n_start:n_end, ]
 
  
  devs <- c()
spl <- (input$cases)
fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 10)
devs <- sd(spl - predict(fit)$y)

temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, i_lim = 6, r_lim = 1)

temp <- as.data.frame(temp)
colnames(temp) <- c("iota", "rho", "Initial infected", "Initial recovered")

pars <- temp %>%  add_column(Method = legend[m]) %>% pivot_longer(!Method, names_to = 'par', values_to = 'value')

result[[m]] <- pars
rm(temp)
rm(pars)
}

# load("SD0.1_SSA-s0_N10000.RData")
seed_effect <- rbind(result[[1]], result[[2]], result[[3]], result[[4]], result[[5]])

ggplot(seed_effect, aes(x = value, color = Method)) + 
  geom_freqpoly(bins = 100) +
  facet_wrap(~ par, scales = "free")  +
  theme_bw() + 
  theme(legend.position = "bottom", aspect.ratio = .618, panel.grid.minor = element_blank()) 

save(seed_effect, file = "seed_varyota.RData")

### It takes less than 3 minutes to arrive here.
```

Varying $\iota$ affects the estimates.

# Population size

## SD0.1 SSA-s0

```{r}
models <- c("SIRDvaryingIota-NormalSD0.1_SSA-s0_N1000",
            "SIRDvaryingIota-NormalSD0.1_SSA-s0_N100000")

result <- vector("list", length(models))

for (m in 1:length(models)){
  selected_country <- models[m] 
   
country <-
  data %>% filter(countriesAndTerritories == selected_country) %>%
  select(dateRep, cases, deaths, countriesAndTerritories, popData2020, infected, recovered, dead) %>%
  mutate(infected = as.numeric(infected), recovered = as.numeric(recovered),
         dead = as.numeric(dead), date_new = dmy(dateRep)) %>% arrange(date_new)

country <- country %>% slice(-1)

seeds <- randomLHS(sims, 4)

seeds <- t(seeds) * c(.5, .1, 6, 1)

#### Fitting

end <- start + 120


  n_start <- which(country$date_new == start)
  n_end <- which(country$date_new == end)
  input <- country[n_start:n_end, ]
 
  
  devs <- c()
spl <- (input$cases)
fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 15)
devs <- sd(spl - predict(fit)$y)

temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, i_lim = 20, r_lim = 10)

temp <- as.data.frame(temp)
colnames(temp) <- c("iota", "rho", "Initial infected", "Initial recovered")

pars <- temp %>% add_column(Method = selected_country) %>% pivot_longer(!Method, names_to = 'par', values_to = 'value')

result[[m]] <- pars
rm(temp)
rm(pars)
}

load("SD0.1_SSA-s0_N10000.RData")
popsize_varyiota <- rbind(pars %>% add_column(popsize = 10000), 
                     result[[1]] %>% add_column(popsize = 1000), 
                     result[[2]] %>% add_column(popsize = 100000))

ggplot(popsize_varyiota, aes(x = value, color = as.factor(popsize))) + 
  geom_freqpoly(bins = 100) +
  facet_wrap(~ par, scales = "free")  +
  theme_bw() + 
  theme(legend.position = "bottom", aspect.ratio = .618, panel.grid.minor = element_blank()) +
  labs(color = "Population size")

save(popsize_varyiota, file = "popsize_varyota.RData")
```

## Data for SD0.1 SSA-s0 N100000

```{r}
selected_country <- "SIRDvaryingIota-NormalSD0.1_SSA-s0_N100000" 
   
country <-
  data %>% filter(countriesAndTerritories == selected_country) %>%
  select(dateRep, cases, deaths, countriesAndTerritories, popData2020, infected, recovered, dead) %>%
  mutate(infected = as.numeric(infected), recovered = as.numeric(recovered),
         dead = as.numeric(dead), date_new = dmy(dateRep)) %>% arrange(date_new)

ggplot(country %>% filter(date_new < dmy("24022020") + 120), aes(x = date_new)) +
  geom_ribbon(aes(ymax = cases, ymin = 0), fill = "darkgreen", alpha = .8) +
  geom_ribbon(aes(ymax = deaths, ymin = 0), fill = "purple", alpha = .8) +
  theme_bw() + theme(panel.grid.minor = element_blank(), aspect.ratio = .618) +
  ylab("Daily cases & deaths") + xlab("Date")
```

Peak is a bit late.

# Intensity of fluctuations

## s0 N10000

```{r fluctuations}

data <- read.csv("artificial220529_varyingIota-Normal.csv", na.strings = "", fileEncoding = "UTF-8-BOM")

models <- c("SIRDvaryingIota-NormalSD0.1_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.09_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.08_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.07_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.06_SSA-s0_N10000",
            "SIRDvaryingIota-NormalSD0.05_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.04_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.03_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.02_SSA-s0_N10000",
            "SIRDvaryingIota-NormalSD0.01_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.009_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.008_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.007_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.006_SSA-s0_N10000",
            "SIRDvaryingIota-NormalSD0.005_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.004_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.003_SSA-s0_N10000",
            # "SIRDvaryingIota-NormalSD0.002_SSA-s0_N10000",
            "SIRDvaryingIota-NormalSD0.001_SSA-s0_N10000")

label <- c(.1, .05, .01, .005, .001)

result <- vector("list", length(models))

for (m in 1:length(models)){
  selected_country <- models[m] 
   
country <-
  data %>% filter(countriesAndTerritories == selected_country) %>%
  select(dateRep, cases, deaths, countriesAndTerritories, popData2020, infected, recovered, dead) %>%
  mutate(infected = as.numeric(infected), recovered = as.numeric(recovered),
         dead = as.numeric(dead), date_new = dmy(dateRep)) %>% arrange(date_new)

country <- country %>% slice(-1)

seeds <- randomLHS(sims, 4)

seeds <- t(seeds) * c(.5, .1, 6, 1)

#### Fitting

end <- start + 120


  n_start <- which(country$date_new == start)
  n_end <- which(country$date_new == end)
  input <- country[n_start:n_end, ]
 
  
  devs <- c()
spl <- (input$cases)
fit <- smooth.spline(x = 1:nrow(input), y = spl, df = 15)
devs <- sd(spl - predict(fit)$y)

temp <- fitting_procedure(seeds = seeds, fun = ll_ode, data = input, stdev = devs, pop = pop, i_lim = 20, r_lim = 10)

temp <- as.data.frame(temp)
colnames(temp) <- c("iota", "rho", "Initial infected", "Initial recovered")

pars <- temp %>% add_column(Method = label[m]) %>% pivot_longer(!Method, names_to = 'par', values_to = 'value')

result[[m]] <- pars
rm(temp)
rm(pars)
}

# load("SD0.1_SSA-s0_N10000.RData")
fluct_varyiota <- rbind(result[[1]], result[[2]], result[[3]],
                        result[[4]], result[[5]])

ggplot(fluct_varyiota, aes(x = value, color = as.factor(Method))) + 
  geom_freqpoly(bins = 100) +
  facet_wrap(~ par, scales = "free")  +
  theme_bw() + 
  theme(legend.position = "bottom", aspect.ratio = .618, panel.grid.minor = element_blank()) +
  labs(color = "Sd")

save(fluct_varyiota, file = "fluct_varyota.RData")
```

It ends up converging to the true values...

# Least squares

## Different seeds

```{r}
data <- read.csv("artificial220517.csv", na.strings = "", fileEncoding = "UTF-8-BOM")

source("least2_function_v1.R")
source("fitting_ls_v1.R")


models <- c("SIRDvaryingIota-NormalSD0.1_SSA-s0_N10000",
  "SIRDvaryingIota-NormalSD0.1_SSA-s1_N10000",
            "SIRDvaryingIota-NormalSD0.1_SSA-s2_N10000",
  "SIRDvaryingIota-NormalSD0.1_SSA-s3_N10000",
  "SIRD_ODE_N10000")

legend <- c("s0", "s1", "s2", "s3", "sd = 0")

result <- vector("list", length(models))

for (m in 1:length(models)){
  selected_country <- models[m] 
   
country <-
  data %>% unique() %>% filter(countriesAndTerritories == selected_country) %>%
  select(dateRep, cases, deaths, countriesAndTerritories, popData2020, infected, recovered, dead) %>%
  mutate(infected = as.numeric(infected), recovered = as.numeric(recovered),
         dead = as.numeric(dead), date_new = dmy(dateRep)) %>% arrange(date_new)


country <- country %>% slice(-1)


seeds <- randomLHS(sims, 4)

seeds <- t(seeds) * c(.5, .1, 6, 1)

#### Fitting

end <- start + 120


  n_start <- which(country$date_new == start)
  n_end <- which(country$date_new == end)
  input <- country[n_start:n_end, ]
 
  

temp <- fitting_procedure(seeds = seeds, fun = ls_ode, data = input, pop = pop, i_lim = 20, r_lim = 5)

temp <- as.data.frame(temp)
colnames(temp) <- c("iota", "rho", "Initial infected", "Initial recovered")

pars <- temp %>% add_column(Method = legend[m]) %>% pivot_longer(!Method, names_to = 'par', values_to = 'value')

result[[m]] <- pars
rm(temp)
rm(pars)
}

seed_effect <- rbind(result[[1]], 
                     result[[2]], result[[3]],
                     result[[4]], result[[5]])

ggplot(seed_effect, aes(x = value, color = Method)) + 
  geom_freqpoly(bins = 100) +
  facet_wrap(~ par, scales = "free")  +
  theme_bw() + 
  theme(legend.position = "bottom", aspect.ratio = .618, panel.grid.minor = element_blank())

save(seed_effect, file = "seed_varyota_ls.RData")
```

## Comparing ls and smoothing

```{r}

load("seed_varyota.RData")
part.a <- seed_effect %>% add_column(Fit = "smoothing")
load("seed_varyota_ls.RData")
part.b <- seed_effect %>% add_column(Fit = "least squares")
ls_logl <- rbind( part.a, part.b)

ggplot(ls_logl, aes(x = value, color = Method)) + 
  geom_freqpoly(aes(linetype = Fit), bins = 100) +
  facet_wrap(~ par, scales = "free")  +
  theme_bw() + 
  theme(legend.position = "bottom", aspect.ratio = .618, panel.grid.minor = element_blank())

# ggplot(seed_effect, aes(x = Method, y = iota/rho, group = Method)) +
#   geom_point()


```

Same results with both fitting procedures.

# End